### 2016-07-11 paper/paper-extra.tex
Lots of stuff

%
% RELATED NOTES (Richard)
%    PersonalJBArchive
%    PSnucleosynthesis/Branch-GalaxyAssemblyAndZDistribution 
%        */bhmetals   : paper.tex, paper-shorter.tex
\documentclass[nofootinbib,twocolumn,prd]{emulateapj}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{color}
\usepackage{ifpdf}
%\usepackage{eps2pdf}

\newcommand\mc{{\cal M}_c{}}
\newcommand\editremark[1]{{\color{red}#1}}
\newcommand\jillianremark[1]{{\color{blue}#1}}
\newcommand\msun{M_\odot}
\newcommand\unit[1]{\text{#1}}
\newcommand\abbrvPSgrb{PSG}
\newcommand\abbrvPSellipticals{PSE}
\newcommand\abbrvKBLowZa{BD2010}
\newcommand\ForInternalReference[1]{}
\newcommand\modelDefault{A}
\newcommand\modelNiino{A'}
\newcommand\modelKW{D}
\newcommand\modelMannucci{D'}
\newcommand\modelKB{B}
\newcommand\modelPanter{C}
\newcommand\cqg{CQG}
\newcommand\E[1]{\left<#1\right>}
%\newcommand\aap{A\&A}
%\newcommand\apss{APSS}
%\newcommand\aaps{AAPS}
%% \newcommand\apjs{ApJ S}
%% \newcommand\aj{AJ}
%% \newcommand\apjl{ApJL}
%% \newcommand\mnras{MNRAS}
%% \newcommand\pasp{PASP}
%% \newcommand\pasj{PASJ}
%% \newcommand\araa{ARAA}
%% \newcommand\physrep{Phys. Rep.}

\begin{document}
\title{Details matter: Similar host galaxies can harbor different merging compact binary populations} 
\author{R.\ O'Shaughnessy}
\affiliation{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, NY 14623, USA}
\email{oshaughn@mail.rit.edu}
\author{ J. Bellovary}
\affiliation{Department of Physics and Astronomy, Vanderbilt University, Nashville TN 37235}
\email{jillian.bellovary@vanderbilt.edu}
\begin{abstract}
% POINT:
% - existence proof by example .. not an attempt to solve the general problem we raise
Galaxy formation
simulations can reproduce present-day galaxies with notably different assembly and star formation histories.  
Recent simulations of compact binary evolution suggest the compact object merger rate depends
sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
assembly can produce more detected compact binaries than typical star formation.   
Using detailed simulations of galaxy and chemical evolution, we demonstrate by example that galaxies with similar present-day appearance can host distinctly different
compact binary populations.
We discuss the implications for transient multimessenger astronomy with compact binary sources. \jillianremark{Add quantifiable stuff when we have it.}
%% Present and future observations of  merging compact binaries can provide valuable constraints on their birthrate
%%  and formation scenarios.  However, recent binary evolution simulations suggest the compact binary merger rate depends
%% sensitively on the progentior's metallicity, to the extent that rare low-metallicity star formation during galaxy
%% assembly can produce more detected compact binaries than typical star formation.    Additionally, galaxy formation
%% simulations can reproduce present-day galaxies with notably different assembly and star formation histories, with
%% corresponding different present-day merger rates.  
%% % POINT 0: Gasoline + popsyn: real work was done
%% Using existing simulations of the assembly,  chemical evolution, and compact-object formation rate of a single halo
%% Milky Way-like galaxy with Gasoline, we construct present-day compact binary merger rates using a simple semianalytic
%% compact binary formation model ($dP/dt \propto Z^\alpha/t $) motivated by detailed calculations in the literature. 
%% % POINT 1: Observations
%% %   - difficult to reconstruct f
%% %   - local particle far away
%% As expected, we show the present-day galactic state can be only weakly correlated with the dominant compact binary
%% fomation rate: the galaxy could form in several ways, including epochs of low-metallicity star formation through major
%% mergers; and present-day mergers rarely occur near gas with their progenitor metallicity, even when not kicked.  
%% %  - but the galaxy 
%% By contrast,  detailed studies which reconstruct the galaxy's star formation history and metallicity evolution could,
%% when stacked, pinpoint the dominant formation process \editremark{need more than handwaving}.
% POINT 2: Revised merger rates
%% Averaging over the cosmological star formation history and galaxy mass-metallicity relation, we demonstrate low
%% metallicity star formation can produce much of the detected population of short GRBs and nearly all gravitatational-wave
%% sources. 
%% % POINT 3: Challenge for observations
%% Therefore,  though a potentially large fraction of binaries formed from low-metallicity gas,  their birth metallicity
%% cannot be identified directly, even with spatially resolved observations. 
%% %
%% We discuss the extent to which observations can distinguish between different scenarios for  low-metallicity compact
%% binary formation.
%
%\textbf{say something using GRBs in paragraph; perhaps mention relative bias in $p(M_g|GRB)$ relative to prior?}
%% \textbf{say something about different SFR histories making present galaxies -- can do more if we can distinguish history
%% of the galaxy}

\end{abstract}

\maketitle

\editremark{We will be very careful to specify how SFR and metallicity are measured, backtracking}

Add me: \cite{2013ApJ...779..102K}
Add me but note conflict \cite{2015MNRAS.453.4337S}, etc

Add me: \citet{2014ARAA..52..415M}

Add me: Hopkins et al  \cite{2016arXiv160508783L} (in context of similar model, building off my prior work)

\section{Introduction}
% POINT: LIGO soon will detect compact binary mergers....
%     ... and possibly EM counter part too
%     ... which will enable new physics
Gravitational waves have been detected from  coalescing  black hole binaries \citep{DetectionPaper,LIGO-O1-BBH}. 
Over the next few years, ground-based gravitational wave detectors like LIGO 
and Virgo  should detect the gravitational
wave signal from many more similar merging compact binaries \citep{LIGO-O1-BBH,RatesPaper,gwastro-EventPopsynPaper-2016}, as well as  binary  neutron stars and black hole-neutron star binaries \citep{LIGO-Inspiral-Rates,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}.
%
The host galaxies of gravitational wave sources will identified, either directly or statistically.  
On the one hand, addition to the strong gravitational wave signal, mergers of neutron stars are expected to  frequently be accompanied by detectable
electromagnetic radiation  via a number of  mechanisms, including a strong but tightly beamed
ultrarelativistic jet and weakly-beamed but more isotropic thermal radiation from an expanding hot shell
\citep[see,e.g.,][and references therein]{2013PhRvL.111r1101C,short-grb-GWCoincidenceEM-MetzgerBerger2011}.  
% People we can't afford the space to cite, alas:
%    - 2015arXiv150101986H   -
%    - 2013MNRAS.430.2121P - Rosswog recent
%    - Shibata group/Koutarou
In addition to providing a wealth of information about the merger event itself, a multimessenger detection will pin down
the sky position and therefore approximate birthplace of each merging binary.    
%
On the other hand, gravitational wave detectors can perform high-precision sky and volumetric localization \cite{2016LRR....19....1A,2016arXiv160307333S}.  As GW detector
networks increase in sensitivity, these localizations will allow statistical and, eventually, unique identification of
host galaxies directly, even without associated electromagnetic emission.
%
As with supernovae and  GRBs, these host galaxy associations   are expected to tightly constrain models for compact binary
formation; see, for comparison,  \citet{2011MNRAS.412.1508M}, \citet{long-grb-GuettaPiran2007},
\citet{2014ARAA..52...43B}, and references therein. %2010ApJ...722.1879M,2010MNRAS.407.1314M}.     
%

% POINT: Huge detail possible
The host galaxies of distant short GRBs have already been extensively investigated, with the associations being used to
draw preliminary conclusions about their progentitors  \cite{2014ARAA..52...43B}.   
%Of necessity, these analyses  on bulk correlations, often stacked to increase significance across multiple sources
%\editremark{clean up/don't piss off   Edo}.  
Unlike GRBs, detected gravitational wave sources will be limited by the range of LIGO to the local universe; for
example,  binary neutron star sources should be closer than $400\unit{Mpc}$.  Due to their proximity, each host galaxy
can be explored at great depth and detail via  position-resolved spectroscopy, enabling detailed position-resolved star-formation
and chemical evolution histories \editremark{citation: need better ones} \citep[see,e.g.][]{2009MNRAS.396..462K,2014MNRAS.444..336C}. \jillianremark{Do you want citations of GRB host observations or just general galaxy stuff?}
% POINT: But is it useful?
In this work, we demonstrate by concrete example that detailed analysis of individual galaxies can be essential to
investigate key physical questions about the origin of compact binary mergers.

This paper is organized as follows. 
In \S \ref{sec:sims}, we describe detailed hydrodynamical simulations of  two galaxies.  Though these two
galaxies are at  present similar to the Milky Way, we
show that their star formation and chemical evolution history have subtle but important differences due to their
distinctive merger histories.   To demonstrate
these differences have a significant practical impact,  in \S \ref{sec:model}, we introduce a simple,
metallicity-dependent  phenomenological model to calculate the present-day rate and mass distribution of compact binary
mergers from  a galaxy's known history.   In \S \ref{sec:results:BBH}, we use this model to demonstrate the binary black hole coalescence rate
depends sensitively on the galaxy's assembly history, in a manner which is far from apparent from its present-day
properties.  
In \S \ref{sec:results} we show that while these two galaxy formation models should produce comparable detectable binary neutron star
populations, these two scenarios could produce dramatically different populations of electromagnetically detectable
merging black hole-neutron star binaries. 
%

%% %   - Delay  time only: 
%% For example, the lag between transients' redshift distribution and the past star formation rate has shown long GRBs are
%% associated with short-lived progenitors (see, e.g., \cite{long-grb-GuettaPiran2007} \cite{grb-long-Virgili2011-TryingEverything} and
%% references therein) and that SNIa merge long after their formation
%% \citep{sn1a-DelayTimeDistribution-Cosmological-Graur2011}. \editremark{other cites}
%% % * \cite{YoungLongGRBHostPredictions2007,FryerGRBProgenitorConstraints2007} : Long bursts.
%% By contrast, short GRBs noticably lag cosmological star formation  \editremark{Berger; }\cite{2007ApJ...665.1220Z,Nakar}

\section{Cosmological simulations produce two distinct Milky-Way-like galaxies}
\label{sec:sims}

\subsection{Simulating galaxy evolution}
To thoroughly study the significance of low metallicity star
formation, we examine cosmological smoothed particle hydrodynamics
(SPH) $N$-body simulations of Milky Way-like galaxies with GASOLINE
\citep{2001PhDT........21S,2004NewA....9..137W}.  These simulations allow us to analyze both
spatially and temporally resolved star formation, and determine not
only the metallicity history of compact object progenitors, but also
their dynamical evolution.  
%maybe move the below section to later and/or consolidate
One simulation we report here, $h258$,
involves the formation of a Milky Way-like disk galaxy whose
progenitors undergo major mergers at $z = 2$ and $z = 0.8$.  This
simulation has been previously discussed by
\citet{2009MNRAS.398..312G,2010ApJ...721L.148B,2011ApJ...742...13B,2011ApJ...728...51B}; however we have
re-run the simulation at a factor of two higher in spatial resolution
and additionally included new physics including a prescription for
cooling by molecular hydrogen \citet{2012MNRAS.425.3058C}.

We selected our simulated region of interest from a 50 Mpc volume of
uniform resolution, and resampled the region at very high
resolution using the volume renormalization technique \citep{1993ApJ...412..455K}.
This technique allows us to follow the detailed physical processes
involved in galaxy evolution in our selected region while still
including large-scale torques from cosmic structure.  We use
WMAP3 cosmology \citep{WMAP} and model the ionizing UV background
with the prescription from \citet{1996ApJ...461...20H}.  Stars form
probabilistically from gas particles which meet density ($n_{min} =
10$ amu cm$^{-3}$) and temperature ($T_{max} = 10^4$ K) thresholds. If a gas
particle meets these criteria, it has a likelihood of forming a star
particle (representing a simple stellar population with a Kroupa IMF
\citep{2001MNRAS.322..231K}) which is given by

\begin{equation}
p = \frac{m_{\rm gas}}{m_{\rm star}} (1 - e^{c^*\rm X_{\rm H_2} \Delta t/t_{\rm form}})
\end{equation}

\noindent
where the star formation efficiency parameter $c^*$ is set to 0.1 such
that our galaxies match the observed Kennicutt-Schmidt law
\citep{1989ApJ...344..685K};  X$_{\rm H_2}$ is the molecular hydrogen fraction of the gas particle; $m_{star}$ and $m_{gas}$ are the star and gas
particle masses;\footnote{In our high-resolution simulations \textbf{ROS-check!}\jillianremark{this number is correct, is that what you want to check? - JMB}, each gas particle starts with a mass
  $m_{gas,0}=26,676 M_\odot$, gaining mass from feedback and losing it to star formation.  Each star particle, when
  formed, has $1/3$ of the progenitor mass of the forming gas particle.  See \cite{2010ApJ...717..121C} for a discussion
of resolution issues in these SPH simulations.} $t_{form}$ is the dynamical time for the gas
particle; and $\Delta t$ is the time between star formation episodes,
which we set to 1 Myr.  We model supernova feedback using the
blastwave formalism described in \citet{1977ApJ...218..148M} and implemented in
our simulations as in \citet{2006MNRAS.373.1074S}.  Each supernova releases
$E_{SN} = 10^{51}$ erg into the surrounding gas with a radius
determined by the blastwave equations.  These particles are not
allowed to cool for the duration of the blastwave, mimicking the
snowplow phase of a supernova explosion.  Previous works have found
that this set of parameters results in realistic galaxies which obey a
number of observed relations such as the mass-metallicity relation
\citep{2007ApJ...655L..17B}, the Tully-Fisher relation \citep{2009MNRAS.398..312G}, and
the size-luminosity relation \citep{2011ApJ...728...51B}, as well as reproduce
the detailed characteristics of bulgeless dwarf galaxies
\citep{2010Natur.463..203G} and the Milky Way \citep{2011ApJ...742...76G}.

Also included in our simulations is a scheme for turbulent metal
diffusion \citep{2010MNRAS.407.1581S}.  Metals are created in supernova explosions
and deposited directly to the gas within the blast radius.  Stellar
masses are converted to ages as described by \citet{1996A&A...315..105R}, and
stars more massive than 8 M$_\odot$ are able to undergo a Type II
supernova.  We follow metal enrichment from both Type II and Type 1a
supernova, with metal yields derived from \citet{1993PhR...227...65W} and
\citet{1986A&A...158...17T} respectively.  From this point onwards metals
diffuse through the surrounding gas, according to 
\begin{equation}
\frac{dZ}{dt}|_{diff} = \nabla (D \nabla {Z}) \\
\end{equation}
%
where the diffusion paremeter $D$ is given by
%
\begin{equation}
D = C_{diff} |S_{ij}| h
\end{equation}
%
and $h$ is the SPH smoothing length, $S_{ij}$ is the trace-free shear
tensor, and $C_{diff}$ is a dimensionless constant which we set to a conservative value of 0.03.

%Sijing finds that C = 0.05 makes for a good comparison to clusters, and that C = 0.05 - 0.1 is expected from turbulence theory.

We identify individual galaxies using the halo finder $AHF$
\citep{2004MNRAS.351..399G,2009ApJS..182..608K}, which identifies halos based on an
overdensity criterion for a flat universe \citep{1997PhDT........14G}.  For this
work, we are focusing on the primary (i.e. most massive) galaxy at any
given redshift, which serves as the progenitor to our $z = 0$ Milky
Way-like galaxy.

\subsection{Two Milky-Way-like galaxies with distinct histories }

{\bf Richard go ahead and edit this however is needed.}  The evolution
of a galaxy, in terms of its stellar mass and metallicity evolution,
depends strongly on its interaction history.  Galaxies which appear
similar at the present day may have had drastically different
histories, which may result in differences in compact object merger
rates.  To investigate whether galaxy history affects the GRB event
rate, we have chosen two simulations which are extremely similar at $z
= 0$ but differ strongly in their merger histories.

The simulation h277 (Boring Galaxy) is a Milky Way analog with a
quiescent merger history.  It experiences its last major merger at $z
= 3$, after which a small number of minor interactions permeate its
life.  This simulation has been shown to emulate several Milky Way
properties, including stellar dynamics
\citep{2012ApJ...758L..23L,2014ApJ...794..151L,2014ApJ...790...89K}, baryon fraction
\citep{2013ApJ...766...56M}, halo properties \citep{2009ApJ...702.1058Z,2010ApJ...721..738Z}, and
disk properties \citep{2011ApJ...728...51B}.  It has a virial mass of $M_{vir} =
6.79 \times 10^{11} \msun$, stellar mass $M_* = 4.24 \times 10^{10}
\msun$, and maximum circular velocity $v_{circ} = 235$ km s$^{-1}$.

  The simulation h258 (Exciting Galaxy), on the other hand, has a much
  more active merger history.  At $z = 1$ there is a 1:1 merger event;
  a gas disk rapidly reforms following the collision (see
  \citet{2009MNRAS.398..312G}), resulting in a massive disk galaxy at $z = 0$
  which looks remarkably similar to the Milky Way and to the other
  simulation, h277.  Prior to the $z = 1$ merger, each of the two
  progenitor galaxies actually experiences its own additional major
  merger events around $z = 2$.  The combination of
  the series of major mergers, plus a number of minor interactions and
  flybys, gives a stark contrast to the relatively quiescent history
  of h277.  At $z = 0$, h258 has a virial mass of $M_{vir} = 7.74
  \times 10^{11} \msun$, stellar mass $M_* = 4.46 \times 10^{10}
  \msun$, and maximum circular velocity $v_{circ} = 242$ km s$^{-1}$.  \editremark{Might we want to show the images of these galaxies?}

  Due to the differences in merger histories, the metallicity
  evolution of h258 and h277 also differ at early times.  Figure
  \ref{fig:TwoGalaxies} shows the history of the mean metallicty of
  recently formed stars vs time (thick solid lines), where we define
  recent as within the past 50 Myr.  The shaded regions correspond to
  one standard devation of the mean, while the thick dashed lines
  represent the top and bottom 90\%.  We see that the early evolution
  of the metal properties of these galaxies does differ - between 0.5
  and 6 Gyr, h277 hosts a {\bf modestly} more metal-rich population than
  h258. \editremark{The difference isn't all that big, I want to address it without sounding foolish.}

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/boring}
\includegraphics[width=\columnwidth]{Figures/exciting}
\caption{\textbf{Two fiducial galaxies}
}
\end{figure*}

\begin{figure*}
\includegraphics[width=\columnwidth]{Figures/sfh}
\includegraphics[width=\columnwidth]{Figures/zvst}
\caption{\label{fig:TwoGalaxies}\textbf{Star formation and metallicity versus time}:  \emph{Left panel}: Star formation history $\dot{M}_{*}$
  versus time.  Black corresponds to galaxy A (``boring'') and red to gaalxy B (``exciting'').  
\emph{Right panel}: A plot of the  metallicity $Z$ of recently-formed stars versus time.  Solid red and black lines show
the mean metallicity; dotted lines correspond to  90\% of newly-born stars have lower metallicity,   or 10\% of
newly-born stars have greater metallicity.  When these two galaxies are $2-3\unit{Gyr}$ old, prior to the first major
merger of  near the peak of their
star-forming history, newly-born stars are created with significantlydifferent metallicity.  Additionally, prior to the
second major merger of galaxy B at $\simeq 6\unit{Gyr}$, stars form in the ``exciting'' galaxy at a systematically lower
metallicity than in the ``boring'' counterpart.
}
\end{figure*}

\section{Model for compact binary formation}
\label{sec:model}

Binary evolution calculations suggest the detection rate depends sensitively on the assumed metallicity, in conjuction
with other parameters \citep[see,\,e.g.][and references
  therein]{popsyn-LowMetallicityImpact-Chris2008,popsyn-LIGO-SFR-2008}.\editremark{New dominik et al}
% POINT: 
%   - final BH spin, mass can change number favored
%   - of course
The detected population of compact binary mergers can further favor low-metallicity environments, e.g. through the total
mass of merging binaries. \jillianremark{The previous sentence is confusing - rephrase?  I have trouble with ``the detected population can further favor low-Z environments'' because that makes no sense.}
% POINT: Simple model to encompass all
To explore plausible binary detection scenarios, we adopt a parameterized formalism for binary evolution and event detection, motivated by
the detailed studies of  \abbrvPSgrb{} and \abbrvPSellipticals.  Specifically, we encapsulate all information about
 binary evolution into two  parameterized response functions.  The funciton ${\cal G}(t| Z) =
dN/dt dM_*$ describes how many mergers per unit initial mass occur after a time $t$ since a starburst,
given ambient metallicity $Z$.  The function $p(m_1,m_2|Z)$ is a metallicity-dependent mass distribution for binaries
formed in a metallicity $Z$.
%
Finally, we account for the substantially increased sensitivity of gravitational wave detectors to binaries of larger
chirp mass $\mc=(m_1 m_2)^{3/5}/(m_1+m_2)^{1/5}$; see \abbrvPSellipticals.
%
In terms of this response function, we can therefore calculate the relative contribution to the LIGO  event rate $r_D$
due to any individual galaxy via the following formula 
\begin{widetext}
\begin{subequations}
\label{eq:RatePerGalaxy}
\begin{eqnarray}
r_D  &\equiv& \int^{0}_{-T_{univ}} dt' d\log Z  \, [\dot{M}_*(t') p(\log Z|t) ] \times \; {\cal G}(0-t' | Z)  
 [\int d\mc p(\mc|Z) (\mc/1.2 M_\odot)^{15/6}]
%\\
%% {\cal R}(z)&\equiv &  \int dx {\cal R} p(z)
\end{eqnarray}
\end{subequations}
\end{widetext}
In this expression, the galactic star formation rate
$\dot{M}_*$ and star formation metallicity distribution $p(\log Z|t)$ are galaxy-dependent and drawn from our
simulations.   \editremark{rewrite this to be average chirp mass at a given metallicity}

We parameterize the response function ${\cal G}(t|Z)$ with two factors: (a) an overall efficiency $\lambda$ for forming
binaries;  and (b) a
cumulative delay time distribution $P_t(<t|Z)$, describing the fraction of binaries born at $t=0$ that merge in less
than time $t$ from a population of binaries $Z$:\footnote{For simplicity, we assume the mass and delay time
  distributions are uncorrelated.  Figure A9 in \abbrvPSgrb{} shows this approximation, while not strictly true, is an
  excellent approximation for merging BH-BH binaries at solar metallicity.  }
\begin{eqnarray}
G(t,x|Z) &=& \lambda(Z) p_x(x|Z) \frac{dP_t}{dt}(t) 
\end{eqnarray}
Previous studies suggest the number of merging compact binaries $\lambda \delta M_*$ produced by a starburst depends
as a power law on metallicity:
\begin{eqnarray}
\label{eq:LambdaVersusZModel}
\lambda(Z) &=& \lambda_o \text{max}[(Z/Z_\odot)^{-a}, F_{max}] 
\end{eqnarray}
with $a\in [0,3]$ and $F_{max}<10^3$
%$\lambda_o F_{max} < 3\times 10^{-4}/M_\odot$
 (see, e.g.\cite{popsyn-LIGO-SFR-2008}, \abbrvKBLowZa).
In this paper we only investigate the dependence of rates with metallicity; we therefore  adopt 
\[
\lambda_0 = 10^{-3}/M_\odot
\]
as a typical value for neutron star compact binaries (see \abbrvPSgrb{}, \abbrvPSellipticals, and \abbrvKBLowZa).
%

For binaries containing neutron stars, for simplicity we adopt the universal delay time distribution
%% Finally, we parameterize the delay time distribution by  either
%% \begin{eqnarray}
%% P_t(<t) &=& \begin{cases}
%% 0 & t<10 \unit{Myr} \\
%% (t/13\unit{Gyr})^{b}& t \in [10\unit{Myr},t_H]
%% \end{cases} 
%% \end{eqnarray}
%% with $b> 0$ or as the special case
\begin{eqnarray}
\frac{dP_t(<t)}{dt} =  \begin{cases}
0 & t<10 \unit{Myr} \\
\frac{1}{t \ln (13 \unit{Gyr}/10\unit{Myr})} & t \in [10\unit{Myr},13\unit{Gyr}]
\end{cases} 
\end{eqnarray}
\abbrvPSgrb{} and \abbrvPSellipticals{} show this distribution is reasonable approximations to compact binary delay
time distributions.\footnote{Simulations suggest the delay time distribution varies from model to model and with mass.
  These variations have less impact on our results than the evolving metallicity distribution of star forming gas.}
%
For suitable choices of parameter, our phenomenological response function is qualitatively consistent with the results of detailed simulations of binary
evolution  \citep{2010ApJ...715L.138B,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2-StarTrackRevised-2012}.

\editremark{Note}:

\section{Formation rate for binary black holes}
\label{sec:results:BBH}

\section{Model for electromagnetic counterparts}
\subsection{Tidal disruption and black hole mass versus metallicity}
At merger, black holes may promptly engulf a companion neutron star \editremark{same papers as for Brandon:Foucart,
  etc}, unless they have sufficiently small mass and sufficiently large spin.  Without  a tidal disruption event prior to merger, the prospects for associated electromagnetic counterparts
are severely diminished.  
%
A neutron star  (NS) coalescing with a black hole (BH) will be tidally disrupted prior to merger if and only if the component
masses, spins, and equation of state permit it.  As described below, only a small fraction of all  possible coalescing
BH-NS binaries have parameters consistent with tidal disruption.  
%
\jillianremark{Grammatical nightmare ahead: }Stellar evolution and hence supernovae remnants are expected to evolve
substantially with birth metallicity, the black hole mass and possibly spin in black hole-neutron star binaries can depend sensitively on binary formation
channels and the fraction of coalescing NS-BH binaries which are tidally disrupted will vary
over cosmic time. 
%
To quantify this effect, we adopt a simple formula for tidal disruption of neutron stars in BH-NS binaries and a simple
phenomenological model for the metallicity-dependent mass distribution of BHs in BH-NS binaries

\begin{figure}
\includegraphics{Figures/fig-mma-TidalDisrupt-MbhA-VersusTheta}
\caption{\label{fig:TidalDisrupt}\textbf{Tidal disruption}:  Following  \citet{2012PhRvD..86l4007F}, a plot of
  the critical spin needed for tidal disruption versus black hole mass and inclination (solid black, dotted black,
  dashed black, solid blue, and dotted blue are $\theta_{LS}=\pi/2,3\pi/8,\pi/4, \pi/8,0$ respectively).  Specifically,
  this figure corresponds to $M_{\rm disk}=0$ in Eq. (\ref{eq:Mdisk}), assuming  a $1.4 M_\odot$ neutron star with compactness ${\cal
    C}=1.66$ in inclined orbit around a spinning black hole of mass $m_{bh}$ and dimensionless spin $\chi_{bh}$.
  Parameters consistent with tidal disruption are above the curves; prompt capture occurs below the
  curve.
}
\end{figure}

% POINT: Trend versus mass and opening angle.  [Be careful: disruption time will depend on opening angle]
To estimate whether a specific compact binary will be tidally disrupted prior to merger, we employ the
following procecure to identify posterior samples consistent with tidal disruption of a neutron star.  First, the
smaller mass must lie within the range of masses allowed by our fiducial NS equation of state (\textbf{MPA1 from Ben}), here
between $0.5 M_\odot$ and $2.5 M_\odot$.  Second, the disruption process must leave behind a remnant disk with nonzero
mass, as estimated using the expression in \cite{2012PhRvD..86l4007F} (his Eqs. (6,12-13)):
\begin{align}
\frac{M_{disk}}{M_{NS}} &=
 0.288 (3 m_{bh}/m_{ns})^{1/3}[1-2 {\cal C}] 
\nonumber \\
&
 -0.148 (m_{bh}/m_{ns})  {\cal C} R_{isco}(a)/m_{bh} 
\label{eq:Mdisk}
\end{align}
where $q=m_{bh}/m_{NS}$ is the binary mass ratio; $a=S_{bh}/m_{bh}^2$ is a dimensionless measure of the black hole spin;
and  where ${\cal C}(m_{ns})=m_{ns}/R_{ns}$ is a mass- and equation-of-state dependent  measure of the neutron star
compactness.  In Foucart's expression, shown above, $R_{isco}(a)$ is the radius of the  innermost stable equitorial circular orbit  of a test particle
about a Kerr black hole  \citep{1972ApJ...178..347B}. % Brandon:  [their Eq. 2.21].   
When the black hole spin is not aligned with the orbit, we use the aligned component of the black hole spin magnitude $a\rightarrow a \cos \theta_{LS}$ in this
expression.   Fixing the neutron star mass $m_{ns}=1.4 M_\odot$ and hence compactness to ${\cal C}=1.66$, this expression implies a two-dimensional surface in
the three-dimensional parameter space $m_{bh},a,\theta_{LS}$ seperates parameters corresponding to tidal disruption to
parameters corresponding to prompt coalescence.
%
After coalescence, the remnant disk generates a tightly beamed electromagnetic counterpart.  For simplicity, we assume
beaming angle is independent of the disruption process or progenitor parameters, so a constant beaming correction factor $f_b$ enters into all
subsequent calculations and can hence be omitted.  
%
Figure \ref{fig:TidalDisrupt} shows contours of this expression for different values of $\theta_{LS}$.  \jillianremark{Add a sentence here describing the significance of the plot.  e.g. low-mass BHs can have a range of spins, regardless of angle, but high mass BHs always get high spins yay.}

\begin{figure}
\includegraphics{Figures/fig-mma-PrecomputeDisrupt}
\caption{\label{fig:TidalDisrupt:Probability}\textbf{Disruption fraction}: Contour plot of the fraction
  $P_{dis}(m_{bh},\chi_{bh})$ of BH-NS binaries with BH mass and spin $m_{bh},\chi_{bh}$ that are expected to tidally
  disrupt using the assumptions in the text.  Contours are shown at $0.1, 1/3, 0.5, 0.9$.}
\end{figure}

% POINT: Distribution of tilt angles
Based on Monte Carlo simulations of supernova kicks as described in the Appendix, we approximate the
cumulative tilt angle distribution $P(<\theta)$ of bound,
coalescing BH-NS binaries by the positive half of a normal distribution
\begin{eqnarray}
P(<\theta)\simeq \int_0^\theta e^{-\theta^2/2\sigma_\theta^2} \frac{2}{\sqrt{2\pi \sigma_\theta^2}} d\theta
\end{eqnarray}
where $\sigma_{\theta}\simeq 1$ is a characteristic angular width that varies relatively weakly with the black hole mass,
 supernova kick velocity, and compact binary formation channel.  This simple distribution is consistent with some detailed population synthesis simulations by
\texttt{StarTrack}.
For simplicity, in this work we adopt $\sigma_{\theta}=1$.   
%  POINT Average over spin distributions
Using this approximate distribution of misalignment angles, Figure \ref{fig:TidalDisrupt:Probability} estimates the fraction
$P_{dis}(m_{bh},\chi_{bh})$ of coalescing binaries with a given black hole mass and spin that should tidally disrupt.  
For simplicity assuming detection weighting does not strongly favor specific relative orientations $\theta$, this
distribution determines the relative likelihood that a gravitational wave event with a given BH mass and spin produces
is consistent with an electromagnetic counterpart.

% POINT: Average over distributions
The relative event rate of electromagnetic counterparts to BH-NS disruptions therefore depends on the intrinsic
distribution $p_d(m_{bh},\chi_{bh})$ of black hole mass and spin among GW-detected BH-NS mergers and the probability
that an event is detected via electromagnetic emission and GW: 
\begin{eqnarray}
p_{em}(m_{bh},\chi_{bh}) = \frac{P_{dis}(m_{bh},\chi_{bh}) p_d(m_{bh},\chi_{bh})}{\int dm_{bh} d\chi_{bh}P_{dis}(m_{bh},\chi_{bh}) p_d(m_{bh},\chi_{bh})}
\end{eqnarray}
where the constant beaming factor cancels.  
%
Because black holes in BH-NS binaries are currently expected to accrete relatively little mass, the spin distribution at merger
should be close to their birth spin distribution.   This distribution is entirely unknown but likely significant, given
tidally-induced spinup prior to merger\editremark{Ott}; we will assume the spin distribution to be uniform between $0.7$
and $1$, independent of formation channel or birth conditions: \editremark{change: make this one Z dependent}
\begin{eqnarray}
p(\chi_{bh}) = \frac{1}{0.3} \theta(\chi_{bh}-0.7)\theta(1-\chi_{bh})
\end{eqnarray}
As discussed in \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012},  the black hole mass distribution depends
sensitively on the birth environment and coalescing binary formation channels; see their Tables 7 and 9 and their Figures 6,9, 11, and 13,
reconstructing the BH mass from the reported chirp masses (red curves).  
Omitting formation channels that predict implausibly high compact binary formation rates (their model A), the
distribution of $m_{bh}$ in BH-NS binaries extends over a wide range.  As a first approximation, the BH mass
distribution in BH-NS binaries does not depend on composition, favoring BH-NS binaries with relatively low BH masses
$m_{bh}\lesssim 5-9 M_\odot$ independent of birth composition.  
% \cite{2012ApJ...749...91F}  : physics of SN explosions used in these mechanisms. Rapid engine needed for mass gap, but
% is there one? 
% mc[m1_, m2_] := (m1 m2)^(3/5)/(m1 + m2)^(1/5)
% Solve[mc[m1, 1.4] == 2.5]
For simplicity, we will adopt a uniform mass distribution between $5$ and $9 M_\odot$:
\begin{eqnarray}
p(m_{bh}) = \frac{1}{4 M_\odot} \theta(m_{bh}-4 M_\odot)\theta(9 M_\odot-m_{bh})
\end{eqnarray}

% POINT: Include delay time: get examples of disruption fraction versus time

\section{Results}
\label{sec:results}

\begin{verbatim}
ACTION ITEMS AND STATUS

* Result script looks good (\texttt{sfr\_to\_rate.py}).  Follow instructions

** Modify script to include tidal disruption options for BHs

** plot 'stanard' results for NS-NS

** set scale to be physical units -- some ad-hoc factors used

*

* Do we get to different answers?  Are they different ENOUGH?

** Overall event rate: different by 20\% between two types. Largely similar.
   - plot versus z_exponent.  Is there a trend

** mass-cut-dependent event rate -- will be more sensitive (e.g., if there
  is a hard cutoff on some masses, as having no EM counterpart)

* Aside: BBH distribution the most sensitive, but not relevant for EM followup
\end{verbatim}

\section{Conclusions}

* BBH properties far from apparent.  No surprise: delays can be long \cite{PSellipticals,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,gwastro-EventPopsynPaper-2016}, particularly because formation in low Z favored,
so system changes significantly from assembly conditions

* EM counterpart properties: delay times are shorter, so more of a surprise that formation scenario matters.  However,
BH properties depend on $Z$, which can influence the mass and spin of BHs that produce EM counterparts.

* Detailed analysis of one host galaxy presented in this work is complementary to the population-based approach used in
\cite{2016arXiv160508783L}, which favored formation in a wide range of galaxies and event times.   % that analysis also
                                % used a poor input from BSE, about the coalescence delay

% POINT: Future directions

\begin{acknowledgements}
ROS acknowledges support from NSF award AST-1412449, via subcontract from the University of Wisconsin-Milwaukee.
%
JB acknolwedges support  from \editremark{writeme}
\end{acknowledgements}

\appendix
\section{Binary evolution and spin-orbit misalignment}

\editremark{ROS to do} -- find python/C code (Davide?) for simple MC; mine old resuts from startrack

%% POINT: Other reference
% 2014arXiv1401.3032W  - fallback details

The progenitors of both components of a BH-NS binary undergo supernova, whose associated asymmetric mass loss and
supernova kicks significantly perturb the orbit  \citep{2013PhRvD..87j4028G,2002MNRAS.329..897H,2000ApJ...541..319K}.   Assuming progeniror stellar spins are
aligned with the orbital plane (e.g., due to strong tides), only asymmetric supernova kicks can change
the orbital plane and introduce misalignment between the BH spin and the orbit.   The distribution of spin-orbit
misalignments expected in any specific binary depends strongly on weakly-constrained binary evolution physics, the
treatment of which is beyond the scope of this article.  Previous detailed calculations have reported on spin-orbit
misalignment \citep{2010ApJ...719L..79F,2008ApJ...682..474B,2004PhRvD..69j2002G}.  That said, relatively simple order-of-magnitude calculations
can explain the typical small misalignment angles expected among merging BH-NS binaries.
% POINT: Small angle tilt forma (follows from simple considerations of angular momentum)
Specifically, in the limit of a small supernova kick $\vec{v}_K$, the orbital plane  change by an angle $\theta \simeq \vec{v}_K \cdot
\hat{z}/v_{orb}$ where $\hat{z}$ is the direction perpendicular to the (circular) orbital plane.  The black hole forms first and cannot
significantly change its spin magnitude or direction through accretion; therefore the two supernovae kicks produced by
forming the black hole and then neutron star add (in quadrature) to produce a characteristic RMS misalignment error
$\E{\theta^2}$.
%
For simplicity, we will adopt the model used in the \texttt{StarTrack} binary evolution code \citep{StarTrack2}: binaries
prior to each supernovae can be approximated as in circular orbits;  supernova kicks on neutron stars produce a Maxwellian
distribution of velocities with 1d velocity dispersion $\sigma_v=265$ \citep{HobbsKicks};  supernova kicks on black holes
$v_{bh}$ are suppressed by fallback, with velocities  $\vec{v}_{K,BH} = (1-f_b)v_{NS}$ suppressed by fallback accretion
onto the black hole; and the dominant formation channels for BH-NS binaries involve common
envelope accretion after the first supernova forms a black hole.   Combining these expressions under these assumptions, we expect the net misalignment  $\theta$ due to both supernova
explosions will have zero mean and variance
\begin{eqnarray}
\label{eq:MisalignmentOrderOfMagnitude}
\E{\theta^2} = \frac{\E{v^2}}{3}\left[ \frac{(1-f_b)^2\epsilon_1}{v_{orb,1}^2}+\frac{\epsilon_2}{v_{orb,2}^2} \right]
\end{eqnarray}
where $v_{orb,1}$ and $v_{orb,2}$ are the orbital velocities at the first and second supernova and where
$\epsilon_1,\epsilon_2$ are factors between 0 and 1 which account for the fact that the average is taken over binaries which \emph{survive} and are not
disrupted by the kick.  Because the second supernova  occurs in a tight orbit (post common
envelope),  the orbital velocity $v_{orb,2}$ at the supernova producing the neutron
star will be  large:  if $M_2$ and $r_2$ are the total mass and orbital separation, $v_{orb,2}=\sqrt{GM_{2}/r}\simeq
440\unit{km/s} \sqrt{M/M_\odot/(r_2/R_\odot)}$, 
% Convert[Sqrt[  GravitationalConstant (mbh SolarMass)/(r SolarRadius)], (Kilo Meter/    Second)]
comparable to or
significantly in excess of the typical supernova kick velocity for massive black holes orbiting the stripped progenitors of
neutron stars at seperations of a few $R_\odot$.  Conversely, at birth the black hole will often capture a significant
amount of the stellar envelope (``fallback'' accretion or prompt collapse) with $f_b\simeq 1$.  Therefore, even though
the initial orbit is wide, orbital  velocity is very small, and hence the orbit can be easily strongly misaligned by
supernova kicks,  the first kick when acting on a massive black hole can produce little to  modest amounts orbital
misalignment.

% POINT: Second term of order unity
For exceptional models of binary evolution, the population of coalescing BH-NS binaries will depend extremely critically on the
a finely-tuned second supernova.  In this case, only a small fraction of surviving BH-NS binaries merge; the BH-NS merger rate will be extremely low and  depend extremely sensitively on
the magnitude of supernova kicks; and the misalignment produced by the second supernova can frequently be very large.  More typically, however, a significant fraction of surviving BH-NS binaries merge
\citep{2008ApJ...675..566O,popsyn-LowMetallicityImpact2-StarTrackRevised-2012,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}; in these cases, the  second supernova's kick does not require fine-tuning to guarantee
coalescence.   In these most common scenarios, the second term in Eq. (\ref{eq:MisalignmentOrderOfMagnitude}) is of
order unity.

% POINT:
As a result, the process of surviving two supernova kicks generally produces a wide  distribution of misalignment
angles concentrated near $\theta\simeq 0$.  This wide distribution is expected to be relatively robust.  Extremely wide
distributions are disfavored kinematically and by the requirement that sufficiently many binaries survive to coalesce.
Conversely, the width of the distribution is bounded below by the contribution from the second supernova kick.\footnote{For simplicity and brevity, we neglect the
  possibility of different modes of supernova explosions and hence kicks, notably via electron capture supernovae \citep{2012ARNPS..62..407J,2004ApJ...612.1044P,2010MNRAS.406..656K,2015ApJ...801...32A}.  That said, small supernova kicks in NS
  birth would further minimize any spin-orbit misaligment and hence further favor the strong dependence of tidal disruption probability on the black hole mass and spin
  distribution described in the text.} 
As a first approximation,  the cumulative distribution of misalignment angles should agree with direct calculations \cite{2010ApJ...719L..79F,2008ApJ...682..474B,2004PhRvD..69j2002G}.

% POINT: 
%% To calculate the necessary inputs to this expression, we adopt one particular family of supernova mechanism:  Fryer et al 
%% \cite{2012ApJ...749...91F}, the model adopted in \texttt{StarTrack} and applied in detailed calculations of binary
%% evolution
%% \cite{popsyn-LowMetallicityImpact2-StarTrackRevised-2012,popsyn-LowMetallicityImpact2b-StarTrackRevised-2013,popsyn-LowMetallicityImpact2c-StarTrackRevised-2014}.
%% Because of significant residual uncertainty in the supernova mechanism, our results are proof-of-concept calculations
%% intended to illustrate plausible effects of composition on remnant masses, spins, and hence misalignments.  Other models
%% for supernovae produce dramatically different initial-final mass relations \cite{2012ApJ...757...69U} and are
%% anticipated to produce equally different different amounts of spin-orbit
%% misalignment in BH-NS binaries.  
%% %
%% Observations of isolated pulsar
%% recoils \cite{1997MNRAS.291..569H,Ar} and binary pulsars \cite{psr-kicks-WongWillemsKalogera2010}
%% strongly constrain the action of random supernova kicks on neutron stars.

\section{Black hole spins in BH-NS binaries}

* References: review in Fragos and McClintock \cite{2015ApJ...800...17F}  and references therein (LMXBs) -- low birth
spins, because of stripped companions; for more on measurement, see \cite{2011CQGra..28k4009M}

* References on BH spin versus mass: angular momentum capture in prompt collapse: O'Connor and Ott and references
therein (note no fallback); note Ugliano has a differen result. See also Clausen et
al (probability of BH vs mass) for more recent
%O'Connor and Ott: a criteria for explosions http://arxiv.org/pdf/1010.5550v2.pdf...note they don't have fallback.
%Clausen et al: trying to find formation probability vs mass http://arxiv.org/pdf/1406.4869v2.pdf (assuming no binaries)
%...note also http://adsabs.harvard.edu/abs/2012ApJ...757...69U (which is complicated)

%* References: \cite{2013ApJ...762..104S}

\section{Outline }
\begin{verbatim}

Intro
  -Important (edo/Brian)
  - More important soon (LIGO; IFU's give detail)
    because these galaxies are local and spatially resolved
    at 400 Mpc distance
  - The question: can we learn about things from detailed host galaxy 
    observations, and if so what?
    Obviously yes (cf. SN Ia, short GRB, etc)
  - Specific questions of interest, particularly for BH-NS origin for short GRBs:
    is there dependence of the event rate and BH mass on the source metallicity

  - Not clear because of kicks, mixing
  - Goal of paper: address this challenge via simulations

Methods
  - JB: specifically target two 'identical' MW galaxies (and their subhalo population)
  - Postproces JB: rhodot; Z of star forming gas
  - Model:
      - predict present-day number (and mass distribution) by formula

Results:
  - Different histories which look the same now can produce different answers
     - key issue: impact of past mergers
  -

Stuff to skip
  - (spin-dependent) tidal disruption threshold: assume all BH-NS disrupt, for simplicity
\end{verbatim}
%1 Kilo Parsec/(400 Mega Parsec)
%Convert[%, Milli ArcSecond] // N

\section{Other discussion}

\section{Implications for transient multimessenger astronomy}

% POINT: Observations provide sample of extragalactic  binaries at merger, and  host galaxies.  Possible to mine large archive of events
% to deduce correlations

On the one hand,  relatively few formation scenarios will correctly reproduce the observed mass and spin distributions
\citep{2004MNRAS.352.1372B,2003ApJ...589L..37B,gwastro-Ilya-ConfProc-NRDA-2010}.  
% POINT: Value of associations
%% Green's function connecting the merger
%% rate $dN/dt$ to the star formation rate host galaxy star formation history and metallicity,
%% both the delay time distribution and correlation any preference for different metallicities can be determined
On the other hand, each host galaxy associates a merger to a unique star formation history and metallicity distribution.
With many events, these associations can potentially determine the ``response function'' for compact binaries: how often star forming gas of a given metallicity evolves into
merging compact binaries;  see, e.g., studies of

%
Highly atypical formation scenarios seemingly should produce events in or near atypical hosts.  
% POINT: Introduce *idea* of comparison: refs for mass-metallicity relation
For example, in the local universe galaxies lie along mass-metallicity  (\cite{2004ApJ...613..898T}; see also
\cite{sfr-ZEvolution-ByGalaxy-Panter2008})
or mass-metallicity-star formation rate  correlations (e.g., \cite{2010MNRAS.408.2115M}).
% POINT: Evolution tricky
[The change in these correlations with redshift is still being investigated; see, e.g., \editremark{citation: Panter;
    Mannucci; }, \cite{2011ApJ...739....1L} and references therein.]
% POINT: previous studies
Based on these correlations, both short  and long GRBs seem to arise from
relatively typical hosts at their redshift; see, e.g., \cite{grb-short-Hosts-Berger2008} and \cite{grb-long-HostMetallicityVsTrend-Mannucci2010}.

%  - Host associations: lots of information
Host galaxy associations provide rich information about a transient's present and past environment.
%
For example, in several cases the metallicity of gas neighboring a long GRB (\citet{ 2008AJ....135.1136M}; \citet{2010AJ....140.1557L}
\editremark{links}), short GRB (\editremark{XXX}), or SN has been directly measured.
%
The precise host offset can be compared to the distribution of light and star formation  \citep{2010ApJ...708....9F}.
%
Finally,  on a host-by-host basis, the delay time between birth and merger has been constrained for long GRBs (\editremark{XXX}), short GRBs \citep{2010ApJ...725.1202L}, and
  SN Ia \citep{2011MNRAS.412.1508M}.
%

Biases towards low-metallicity star formation have been extensively studied in the context of long GRBs and their host galaxies.
%
For example, \cite{2009ApJ...702..377K} demonstrated that a sufficiently strong bias towards low-metallicity star formation would predict
most  events in the local universe occur low-mass and dwarf galaxies.
%<
For less extreme metallicity biases, subsequent calculations by  \citet{2011MNRAS.417..567N} demonstrated that the
metallicity distribution within galaxies will usually lead to events in a wide range of host galaxies in the local universe.

\subsection{Local metallicity measurements?}
To this point, we have emphasized \emph{global} measurements of the galactic star formation history.  With the advent of
IFUs and position-resolved spectroscopy, observers can now probe the star formation history and metallicity of
individual gas packets at different points in a galaxy.  
%
These highly-detailed probes will be essential tools to develop a comprehensive model of the galaxy's merger and
chemical evolution history.  That said, the metallicity of stars and gas adjacent to a specific merger event provides
few direct, unambiguous clues to a compact binary merger's progenitors.  
On the one hand, compact binaries are kicked by supernovae, moving substantially away from their birth position
%  2014arXiv1407.3796S : implications for nucleosynthesis
\cite{2013ApJ...776...18F,2014ARAA..52...43B}. 
On the other hand, particularly during the epoch of galaxy formation, galaxies are well-mixed: stars and adjacent gas generally do not have
similar chemical composition.  
% POINT: Host tracer gas --we are optimistically assuming the metallicity 
%  can be measured.  A seperate question is whether you can measure Z in the local environment at all.
These mixing effects have been previously recognized as an obstacle to interpreting transient event spectra.  For
example,   \cite{2010MNRAS.402.1523P} previously demonstrated that absorbing gas neighboring transient events (there, long GRBs)
would generally have high metallicity, even for low-metallicity progenitors.   
%
\citet{2010MNRAS.402.1523P} have previously used hydrodynamical simulations to demonstrate that observed ambient
metallicities (there, using damped Lyman $\alpha$ absorbers in the host) do not tightly constrain the metallicity
distribution of the progenitor; see, e.g., their Fig 3.
%

\bibliography{bhmetals}
\end{document}